---
title: "Lab Solutions"
author: "Bernado Bernal, Tyler Chung, Marco DiPierro"
format:
  html: 
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---
## Functions and Dataframes

The goal of this lab is to get a solid handle on how to develop functions in hand with a dataframe. As doing so can greatly decrease the amount of repeated code needed to get similar output from looking through a dataframe multiple times at different points of time. Think about how you might have a dataframe of candy and you end with a table of chocolate candies with selected information, only to find out you have to make a similar table with hard candies instead. To make the same table twice but with one difference here would be a lot of repeated code. Luckily with a function that can be reduced to a single line in both cases along with the code to make the function.

The dataset we are using for this lab comes from [kaggle](https://www.kaggle.com/datasets/amanbarthwal/imdb-movies-data). The dataset has various information on 10,000 movies collected from *IMDB*. As such you can expect to find title, genre, and rating information from this dataset, along with some other information. Visit the link to get the whole data dictionary in order to find out what each column means, we will be using version two of the data. While the data is not overly complex it contains enough information to make a couple of functions that pull out or modify data in interesting ways.

## Setup
```{r}
#load in libraries
library(tidyverse)
library(knitr)
```

```{r}
#load in imdb-movies-dataset.csv
movies <- read_csv("imdb-movies-dataset.csv")

```


## Data Cleaning

**1. Cleaning up our data to work with**

Before we can begin working with this dataset, we need to clean it up. Looking at the movies dataset you may notice some missing values, duplicate entries, and columns that won't be useful for our analysis.
Using one dplyr pipeline, clean the movies dataset by:

Removing any rows where Title, Genre, or Rating are missing
Removing any duplicate movies based on Title
Removing columns that will not be used Poster, Review Title, Review, and Description

Assign your cleaned data to a new variable called movies_clean use this data going forward.
```{r}
movies_clean <- movies |>
  filter(
    !is.na(Title),
    !is.na(Genre),
    !is.na(Rating)
  ) |>
  distinct(Title, .keep_all = TRUE) |>
  rename(Duration = "Duration (min)") |>
  select(Title, Year, Certificate, Duration, Genre, Rating,
    Cast, Director, Metascore, Votes)
```

## Simple Functions

Lets get into functions with a simple one

**2. Write a function to filter a table by a specific actor or genre**

Looking at the movies dataset it could be seen that both genre and cast contain a list of comma separated values, rather than only containing one value. For this function if we are given a specific actor or genre we should return the rows that contain that value, but with only the specified value given in the respective column. If given Horror as a genre then we should only return rows that contain Horror as a genre and mutate them to only contain Horror as a genre rather than something like "Horror, Comedy".

To help guide you in making this function you are taking 3 inputs.

1. A dataframe
2. A given column
3. A character value 

:::{.callout-tip}
## Working with dplyr in functions

Remember that when in a function it would be useful to use `rlang` in order to get tidy evaluation for a column passed to a function.

:::

```{r}
#code your function
single_actor_genre <- function(df, col, str) {
  df |> 
    filter(str_detect(string = {{col}},
                      pattern = str)) |> 
    mutate({{col}} := str_extract(string = {{col}},
                                 pattern = str))
}
```

Check your function
```{r}
#will likely have to change function name and input here to see if your code worked
single_actor_genre(movies_clean, Genre, "Horror") |> 
  select(Title, Genre) |> 
  head(5) |> 
  kable()
```

**3. Adding input validation to our function**
```{r}

single_actor_genre <- function(df, col, str) {
  stopifnot("`df` must be a dataframe." = is.data.frame(df))
  stopifnot("`str` must be a character string." = is.character(str))
  
  col_name <- rlang::as_name(rlang::ensym(col))

  stopifnot(
    "`col` must be a column in df." = col_name %in% names(df)
  )
  
  df |> 
    filter(str_detect(string = {{col}},
                      pattern = str)) |> 
    mutate({{col}} := str_extract(string = {{col}},
                                 pattern = str))
}

```

## More complex functions

Now that we have added input validation to our simple function we should be able to take a step up in making a function and add some default values to a function. Which just means that if a value isn't passed it will default to a preset value.

**4.Create a function that outputs yearly information related to a genre or actor**

The goal of this function is simple, given a genre or actor we want to collapse information into a single year. For example if there where 6 horror movies in 1966 we simply want to take the average or sum of some values into a single year. Here is a list of input variables and output columns to help you out in creating this function.

**input**  

1. a dataframe
2. a column that's either Genre or Cast
3. a character value that relates to genre or cast
4. a good rating value that should default to 7 if nothing is passed to it
5. an average rating value that should default to 4 if nothing is passed to it

Output columns

1. column passed 
2. year: should not have any repeated years
3. avg_rating: average rating from that year
4. total_votes: total votes from that year
5. quality: contains 3 possible character values
    - good: if avg_rating was above the good rating value
    - average: if avg_rating was above average rating value but below the good rating value
    - bad: if avg_rating is below the average rating value
  
Remember to add input validation here as its good practice and stops any unexpected error before they happen.

```{r}

yearly_info <- function(df, col, str, good_rating = 7, average_rating = 4) {
  #input validation
  stopifnot("`df` must be a dataframe." = is.data.frame(df))
  stopifnot("`str` must be a character string." = is.character(str))
  stopifnot("`col` must be `Genre` or `Cast`." = str_detect(string = deparse(substitute(col)),pattern = "Genre|Cast"))
  
  #function
  df |> 
  single_actor_genre({{col}}, str) |> 
  group_by({{col}}, Year) |> 
  summarise(avg_rating = mean(Rating),
            total_votes = sum(Votes)) |> 
  mutate(quality = case_when(avg_rating > good_rating ~ "good",
                             avg_rating > average_rating ~ "average",
                             .default = "bad"))
}

```

**Test your function here**
```{r}
#mess around with this code to make sure you getting error when providing things that arnt expected
yearly_info(movies_clean, Genre, "Horror") |> 
  head(5) |> 
  kable()
```

## The power of functions

As you might have noticed when writing your more complex function the true beauty of functions is how little you have to repeat code and how they just shorten everything. To get a more solid appreciation of them we will now use our more complex function in getting some fun data

**5. Make a table with yearly information on 2 or more genres or actors**

Using the more complex function you made, save a dataframe that contains yearly information on 2 or more genres or actors but you must stick to your chosen category. As in only 2 actors if you pick actors and not 1 actor or one genre. It should also contain the title of the highest rated movie of that year, its duration, its rating and votes.You can also drop cast or genre depending on what you pick to look into. For example if you pick 2 actors then you can drop genre, but if you pick 2 genres you can drop cast.

:::{.callout-tip}
## Expanding rows in a dataframe

It would be useful to look into functions that could help you **bind** rows together or a **map function that returns a dataframe**

:::

```{r}
genre <- map_dfr(c("Horror", "Comedy", "Action"),~ yearly_info(movies_clean, Genre, .x)) |> 
  inner_join(movies_clean, by = c("Genre", "Year")) |> 
  group_by(Genre, Year) |> 
  slice_max(Rating, n = 1, with_ties = FALSE) |> 
  select(-Metascore, -Director, -Certificate, -Cast) |> 
  ungroup()
```
Take a moment to think about how many more lines of code would be needed if we didn't already have functions to handle a lot of the busy work and appreciate your work so far.  

## References

Markdown syntax guide used for links and list  
<https://www.markdownguide.org/basic-syntax/>

Quarto guide in including callout blocks  
<https://quarto.org/docs/authoring/callouts.html>

Getting a variable name as a character  
<https://stackoverflow.com/questions/14577412/how-to-convert-variable-object-name-into-string>

Function that returns a dataframe when mapping  
<https://purrr.tidyverse.org/reference/map_dfr.html>

