---
title: "Lab 9: Functions and Dataframes"
format:
  html: 
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

## Functions and Dataframes

The goal of this lab is to get a solid handle on how to develop functions in hand with a dataframe. As doing so can greatly decrease the amount of repeated code needed to get similar output from looking through a dataframe multiple times at different points of time. Think about how you might have a dataframe of candy and you end with a table of chocolate candies with selected information, only to find out you have to make a similar table with hard candies instead. To make the same table twice but with one difference here would be a lot of repeated code. Luckily with a function that can be reduced to a single line in both cases along with the code to make the function.

The dataset we are using for this lab comes from [kaggle](https://www.kaggle.com/datasets/amanbarthwal/imdb-movies-data). The dataset has various information on 10,000 movies collected from *IMDB*. As such you can expect to find title, genre, and rating information from this dataset, along with some other information. Visit the link to get the whole data dictionary in order to find out what each column means, we will be using version two of the data. While the data is not overly complex it contains enough information to make a couple of functions that pull out or modify data in interesting ways.

## Setup

```{r}
#load in libraries
library(tidyverse)
library(knitr)
```

```{r}
#load in imdb-movies-dataset.csv
movies <- read_csv("imdb-movies-dataset.csv")
```

---

## Data Cleaning

**1. Clean the dataset before analysis**

Before we can begin working with this dataset, we need to tidy it up. The raw data contains missing values, duplicate entries, and columns we won't need.

Using a single `dplyr` pipeline, create a cleaned dataset called `movies_clean` by:

- Removing any rows where `Title`, `Genre`, or `Rating` are missing
- Removing duplicate movies, keeping only the first occurrence of each `Title`
- Renaming `Duration (min)` to `Duration` (the space and parentheses will cause issues later)
- Keeping only the columns: `Title`, `Year`, `Certificate`, `Duration`, `Genre`, `Rating`, `Cast`, `Director`, `Metascore`, `Votes`

Use `movies_clean` for all questions going forward.

```{r}
#Q1 Code

```

## Simple Functions

Lets get into functions with a simple one

**2. Write a function to filter a table by a specific actor or genre**

Looking at the movies dataset it could be seen that both genre and cast contain a list of comma separated values, rather than only containing one value. For this function if we are given a specific actor or genre we should return the rows that contain that value, but with only the specified value given in the respective column. If given Horror as a genre then we should only return rows that contain Horror as a genre and mutate them to only contain Horror as a genre rather than something like "Horror, Comedy".

To help guide you in making this function you are taking 3 inputs.

1. A dataframe
2. A given column
3. A character value 

```{r}
#Q2 Code

```

Test your function

```{r}
#function test
#will likely have to change function name and input here to see if your code worked
single_actor_genre(movies_clean, Genre, "Horror") |> 
  select(Title, Genre) |> 
  head(5) |> 
  kable()
```

**3. Add input validation to your function**

Rewrite the function to include checks that guard against bad inputs. Specifically, your function should stop with a clear error message if:

- `df` is not a dataframe
- `str` is not a character string
- The column passed as `col` does not exist in `df`

```{r}
#Q3 Code

```

## More complex functions

A big part of functions is that they actually use other functions to help out, this is often called a helper function. Since we already made a function to help us in filtering the table out we should make use of it as a helper function to make a more complex function.

While we are at it we should also add some default values to a function. Which just means that if a value isn't passed it will default to a preset value.

**4.Create a function that outputs yearly information related to a genre or actor**

The function we will now be making will collapse information that is filtered out from the previous function into rows based on year. For example if the previous function was used to filter horror genre movies, and it returned us 6 movies from 2011 we will want collapse the information into one row of 2011. To do this you would want to take the average rating and sum of votes.

While at it we should also introduce some default values to say if that year had a good, average, or bad quality of movies. So for this part if a year had an average rating above a 7 it should be considered good. Anything between a 4 and 7, including the 7 should be average, and anything below that should be bad. Although a user should be able to change those 7 and 4 values if they want

Here is an example of what the output should look like

| Genre | Year | avg_rating | total_votes| quality |
|:-----:|:----:|:----------:|:----------:|:-------:|
| Horror| 2011 |    7.1     | 104644     | good    |
| Horror| 2010 |    6.2     | 529576     | average |
| Horror| 2009 |    4       | 92455      | bad     |

: possible output

The input of this function should be straightforward, since your using a helper function you need its input to be included in this function as well. Then you are just adding the default variables that where mentioned before.

```{r}
#Q4 Code

```

Test your function

```{r}
#function test
#will likely have to change function name and input here to see if your code worked
yearly_info(movies_clean, Genre, "Horror") |> 
  head(5) |> 
  kable()
```

---

## The power of functions

As you might have noticed when writing your more complex function the true beauty of functions is how little you have to repeat code and how they just shorten everything. To get a more solid appreciation of them we will now use our more complex function in getting some fun data

**5. Make a table with yearly information on 3 genres**

Using the more complex function you made, save a dataframe that contains yearly information on Horror, Comedy, and Action. It should also contain the title of the highest rated movie of that year for each genre, its duration, its rating and votes.

:::{.callout-tip}
## Expanding rows in a dataframe

It would be useful to look into functions that could help you **bind** rows together or something like `map_dfr`

:::

```{r}
#Q5 Code

```

Take a moment to appreciate how little repeated code was needed here compared to doing this without functions.

## Presenting Data Nicely

**6. Make a polished summary table**

Using the dataframe you created in question 5, create a clean, readable table. To keep things manageable, only include rows from **2010 onward**, filter out any rows where `Year` is missing, and show only the top 10 rows by average rating.

Your table should:

- Show only the most relevant columns (genre, year, summary stats, best movie info)
- Have clean, human-readable column names
- Round any numeric values to two decimal places
- Include a descriptive caption about the table

```{r}
#Q6 Code

```

**7. Visualize average rating over time by genre**

Using the dataframe you created in question 5, create a clean readable line graph with `ggplot2` showing how the average yearly rating for each genre has changed over time.

Your plot should:

- Use `color` to distinguish genres
- Use `shape` on the points to reflect the `quality` label (`"good"`, `"average"`, `"bad"`)
- Have a clear title, axis labels, and legend
- Use a clean theme

```{r}
#Q7 Code

```

## Statistical Analysis

**8a. Using the same dataset, let's see whether there's a correlation between a movie's duration and its rating. Create a scatterplot of `Duration` vs `Rating` and include a linear regression line. Describe the scatterplot.**

:::{.callout-tip}
Don't worry about labels for this question.
:::

```{r}
#Q8a Code

```

**8b. Fit a simple linear regression model predicting Rating by Duration. Report the R-squared value and give a brief interpretation.**

```{r}
#Q8b Code

```


**8c. Now nest your regression by genre, and report which genre had the highest R-squared. What are possible limitations of the model we used? (hint: look at the number of rows for each genre!)**

```{r}
#Q8c Code

```

