---
title: "Lab Solutions"
author: "Bernado Bernal, Tyler Chung, Marco DiPierro"
format:
  html: 
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---
## Functions and Dataframes

The goal of this lab is to get a solid handle on how to develop functions in hand with a dataframe. As doing so can greatly decrease the amount of repeated code needed to get similar output from looking through a dataframe multiple times at different points of time. Think about how you might have a dataframe of candy and you end with a table of chocolate candies with selected information, only to find out you have to make a similar table with hard candies instead. To make the same table twice but with one difference here would be a lot of repeated code. Luckily with a function that can be reduced to a single line in both cases along with the code to make the function.

The dataset we are using for this lab comes from [kaggle](https://www.kaggle.com/datasets/amanbarthwal/imdb-movies-data). The dataset has various information on 10,000 movies collected from *IMDB*. As such you can expect to find title, genre, and rating information from this dataset, along with some other information. Visit the link to get the whole data dictionary in order to find out what each column means, we will be using version two of the data. While the data is not overly complex it contains enough information to make a couple of functions that pull out or modify data in interesting ways.

## Setup
```{r}
#load in libraries
library(tidyverse)
library(knitr)
```

```{r}
#load in imdb-movies-dataset.csv
movies <- read_csv("imdb-movies-dataset.csv")

```


## Data Cleaning

**1. Cleaning up our data to work with**

Before we can begin working with this dataset, we need to clean it up. Looking at the movies dataset you may notice some missing values, duplicate entries, and columns that won't be useful for our analysis.
Using one dplyr pipeline, clean the movies dataset by:

Removing any rows where Title, Genre, or Rating are missing
Removing any duplicate movies based on Title
Removing columns that will not be used Poster, Review Title, Review, and Description

Assign your cleaned data to a new variable called movies_clean use this data going forward.
```{r}
movies_clean <- movies |>
  filter(
    !is.na(Title),
    !is.na(Genre),
    !is.na(Rating)
  ) |>
  distinct(Title, .keep_all = TRUE) |>
  rename(Duration = "Duration (min)") |>
  select(Title, Year, Certificate, Duration, Genre, Rating,
    Cast, Director, Metascore, Votes)
```

## Simple Functions

Lets get into functions with a simple one

**2. Write a function to filter a table by a specific actor or genre**

Looking at the movies dataset it could be seen that both genre and cast contain a list of comma separated values, rather than only containing one value. For this function if we are given a specific actor or genre we should return the rows that contain that value, but with only the specified value given in the respective column. If given Horror as a genre then we should only return rows that contain Horror as a genre and mutate them to only contain Horror as a genre rather than something like "Horror, Comedy".

To help guide you in making this function you are taking 3 inputs.

1. A dataframe
2. A given column
3. A character value 

```{r}
#code your function
single_actor_genre <- function(df, col, str) {
  df |> 
    filter(str_detect(string = {{col}},
                      pattern = str)) |> 
    mutate({{col}} := str_extract(string = {{col}},
                                 pattern = str))
}
```

Check your function
```{r}
#will likely have to change function name and input here to see if your code worked
single_actor_genre(movies_clean, Genre, "Horror") |> 
  select(Title, Genre) |> 
  head(5) |> 
  kable()
```

**3. Add input validation to the function you wrote in (2). You can copy and paste your code.**
```{r}

single_actor_genre <- function(df, col, str) {
  stopifnot("`df` must be a dataframe." = is.data.frame(df))
  stopifnot("`str` must be a character string." = is.character(str))
  
  col_name <- rlang::as_name(rlang::ensym(col))

  stopifnot(
    "`col` must be a column in df." = col_name %in% names(df)
  )
  
  df |> 
    filter(str_detect(string = {{col}},
                      pattern = str)) |> 
    mutate({{col}} := str_extract(string = {{col}},
                                 pattern = str))
}

```

## More complex functions

A big part of functions is that they actually use other functions to help out, this is often called a helper function. Since we already made a function to help us in filtering the table out we should make use of it as a helper function to make a more complex function.

While we are at it we should also add some default values to a function. Which just means that if a value isn't passed it will default to a preset value.

**4.Create a function that outputs yearly information related to a genre or actor**

The function we will now be making will collapse information that is filtered out from the previous function into rows based on year. For example if the previous function was used to filter horror genre movies, and it returned us 6 movies from 2011 we will want collapse the information into one row of 2011. To do this you would want to take the average rating and sum of votes.

While at it we should also introduce some default values to say if that year had a good, average, or bad quality of movies. So for this part if a year had an average rating above a 7 it should be considered good. Anything between a 4 and 7, including the 7 should be average, and anything below that should be bad. Although a user should be able to change those 7 and 4 values if they want

Here is an example of what the output should look like

| Genre | Year | avg_rating | total_votes| quality |
|:-----:|:----:|:----------:|:----------:|:-------:|
| Horror| 2011 |    7.1     | 104644     | good    |
| Horror| 2010 |    6.2     | 529576     | average |
| Horror| 2009 |    4       | 92455      | bad     |

: possible output

The input of this function should be straightforward, since your using a helper function you need its input to be included in this function as well. Then you are just adding the default variables that where mentioned before.

```{r}

yearly_info <- function(df, col, str, good_rating = 7, average_rating = 4) {
  #input validation
  stopifnot("`df` must be a dataframe." = is.data.frame(df))
  stopifnot("`str` must be a character string." = is.character(str))
  stopifnot("`col` must be `Genre` or `Cast`." = str_detect(string = deparse(substitute(col)),pattern = "Genre|Cast"))
  
  #function
  df |> 
  single_actor_genre({{col}}, str) |> 
  group_by({{col}}, Year) |> 
  summarise(avg_rating = mean(Rating),
            total_votes = sum(Votes)) |> 
  mutate(quality = case_when(avg_rating > good_rating ~ "good",
                             avg_rating > average_rating ~ "average",
                             .default = "bad"))
}

```

**Test your function here**
```{r}
#mess around with this code to make sure you getting error when providing things that arnt expected
yearly_info(movies_clean, Genre, "Horror") |> 
  head(5) |> 
  kable()
```

## The power of functions

As you might have noticed when writing your more complex function the true beauty of functions is how little you have to repeat code and how they just shorten everything. To get a more solid appreciation of them we will now use our more complex function in getting some fun data

**5. Make a table with yearly information on 3 genres**

Using the more complex function you made, save a dataframe that contains yearly information on 3 genres. It should also contain the title of the highest rated movie of that year for each genre, its duration, its rating and votes.

```{r}
three_genres <- map_dfr(c("Horror", "Comedy", "Action"),~ yearly_info(movies_clean, Genre, .x)) |> 
  inner_join(movies_clean, by = c("Genre", "Year")) |> 
  group_by(Genre, Year) |> 
  slice_max(Rating, n = 1, with_ties = FALSE) |> 
  select(-Metascore, -Director, -Certificate, -Cast) |> 
  ungroup()
```

Take a moment to think about how many more lines of code would be needed if we didn't already have functions to handle a lot of the busy work and appreciate your work so far.  

## Presenting Data Nicely

**6. Create a nicer looking table for the 3 genres**

Using the dataframe from question 5, create a clean readable table with `select()`, `rename()`, `arrange()`, `mutate()`, and `kable()`. Clean up the column names, round numeric values, and arrange by genre and year.


```{r}
three_genres |>
  filter(Year >= 2010, !is.na(Year)) |>
  select(Genre, Year, avg_rating, total_votes, quality, Title, Duration, Rating) |>
  arrange(Genre, Year) |>
  mutate(avg_rating = round(avg_rating, 2)) |>
  slice_max(avg_rating, n = 10) |>
  kable(
    caption   = "Yearly Film Performance by Genre (2010–present): Horror, Comedy, and Action",
    col.names = c("Genre", "Year", "Avg Rating", "Total Votes",
                  "Quality", "Best Movie", "Duration (min)", "Movie Rating")
  )
```

---

**7. Create a graph based on the 3 genres**

Using the dataframe from 5 and 6, create a graph with `ggplot2` that shows how the average yearly rating for each genre has changed over time. Use `geom_line()` and `geom_point()` to show the trend and `color` to distinguish the genres. Include clear axis labels and a title.

```{r}
#| fig-width: 10
#| fig-height: 5
three_genres |>
  filter(!is.na(Year)) |>
  select(Genre, Year, avg_rating, quality) |>
  distinct() |>
  ggplot(aes(x = Year, y = avg_rating, color = Genre, group = Genre)) +
  geom_line() +
  geom_point(aes(shape = quality), size = 3) +
  labs(
    title   = "Average IMDB Rating by Year and Genre",
    x       = "Year",
    y       = "Average Rating",
    color   = "Genre",
    caption = "Source: IMDB Movies Dataset (Kaggle)"
  ) +
  theme_bw()
```

## Statistical Analysis

**8a. Using the same dataset, let's see whether there's a correlation between a movie's duration and its rating. Create a scatterplot of `Duration` vs `Rating` and include a linear regression line. Describe the scatterplot.**


```{r}

three_genres |> 
  ggplot(aes(x = Duration, y = Rating)) +
  geom_point(alpha = .8) +
  geom_smooth(method = "lm") 

```
There is high variability when it comes to rating. Duration is mostly clustered around 100 minutes, so the relationship doesn't seem to be completely linear. The regression line is slightly upward sloping, suggesting a weak positive association between duration and rating could exist.

**8b. Fit a simple linear regression model predicting Rating by Duration. Report the R-squared value and give a brief interpretation.**

```{r}

model <- lm(Rating ~ Duration, data = three_genres)

summary(model)

```
The (multiple) R-squared value for this regression is 0.03746. This means that only 3.746% of the variance in rating can be explained by duration, which is very low.

**8c. Now nest your regression by genre, and report which genre had the highest R-squared. What are possible limitations of the model we used? (hint: look at the number of rows for each genre!)**

```{r}
by_genre <- three_genres |> 
  nest(genre_dat = -Genre) |> 
  mutate(model2 = map(genre_dat, 
                   ~ lm(Rating ~ Duration, data = .x)),
         stats = map(model2, broom::glance)) |> 
  select(Genre, stats) |>
  unnest(stats) |>
  select(Genre, r.squared)

by_genre |> 
  slice_max(r.squared) |> 
    kable()
```

The 'Action' genre has the highest R-squared, at 0.2657681. A limitation of this model is the very small sample size, especially for each individual genre. The genre overall dataframe has only 116 observations, the "Action" genre only has 13 observations.

## References

Markdown syntax guide used for links and list  
<https://www.markdownguide.org/basic-syntax/>

Quarto guide in including callout blocks  
<https://quarto.org/docs/authoring/callouts.html>

Getting a variable name as a character  
<https://stackoverflow.com/questions/14577412/how-to-convert-variable-object-name-into-string>

Function that returns a dataframe when mapping  
<https://purrr.tidyverse.org/reference/map_dfr.html>

