---
title: "Lab 9: Functions and Dataframes"
format:
  html: 
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

## Functions and Dataframes

The goal of this lab is to develop a solid handle on how to build functions that work alongside a dataframe. Functions are powerful because they eliminate repeated code. Imagine you have a dataframe of candy and you need a table showing only the chocolate candies — then your boss asks for the same table but for hard candies instead. Without a function, you'd copy and paste nearly identical code twice. With a function, each table is just a single line of code.

The dataset we are using comes from [Kaggle](https://www.kaggle.com/datasets/amanbarthwal/imdb-movies-data). It contains information on 10,000 movies collected from *IMDB*, including title, genre, rating, cast, and more. Visit the link for the full data dictionary. We will be using version two of the data.

## Setup

```{r}
library(tidyverse)
library(knitr)
```

```{r}
movies <- read_csv("imdb-movies-dataset.csv")
```

---

## Data Cleaning

**1. Clean the dataset before analysis**

Before we can begin working with this dataset, we need to tidy it up. The raw data contains missing values, duplicate entries, and columns we won't need.

Using a single `dplyr` pipeline, create a cleaned dataset called `movies_clean` by:

- Removing any rows where `Title`, `Genre`, or `Rating` are missing
- Removing duplicate movies, keeping only the first occurrence of each `Title`
- Renaming `Duration (min)` to `Duration` (the space and parentheses will cause issues later)
- Keeping only the columns: `Title`, `Year`, `Certificate`, `Duration`, `Genre`, `Rating`, `Cast`, `Director`, `Metascore`, `Votes`

Use `movies_clean` for all questions going forward.

```{r}
# your code here
```

---

## Simple Functions

**2. Write a function to filter a table by a specific actor or genre**

Both the `Genre` and `Cast` columns contain comma-separated lists of values (e.g., `"Horror, Comedy"` or `"Tom Hanks, Robin Wright"`). Write a function called `single_actor_genre` that:

1. Filters the dataframe to only rows where a given column **contains** a specified string
2. Mutates that column so it shows **only** the matched value (not the full comma-separated list)

For example, passing `"Horror"` as the genre should return rows that include Horror as a genre, with the `Genre` column showing only `"Horror"`.

Your function should accept three arguments:

1. `df` — a dataframe
2. `col` — the column to search (e.g., `Genre` or `Cast`)
3. `str` — the character value to search for

:::{.callout-tip}
## Working with dplyr inside functions

When passing a column name into a function, use `{{ col }}` (curly-curly) for `dplyr` verbs. This is part of `rlang`'s tidy evaluation system.
:::

```{r}
# your function here
```

Check your function:

```{r}
single_actor_genre(movies_clean, Genre, "Horror") |> 
  select(Title, Genre) |> 
  head(5) |> 
  kable()
```

---

**3. Add input validation to your function**

Rewrite `single_actor_genre` to include checks that guard against bad inputs. Specifically, your function should stop with a clear error message if:

- `df` is not a dataframe
- `str` is not a character string
- The column passed as `col` does not exist in `df`

:::{.callout-tip}
## Stopping with a message

`stopifnot()` is a convenient way to validate inputs. You can supply a named argument where the name becomes the error message:

```r
stopifnot("my message" = some_condition)
```

To check whether a `col` argument (passed with `{{ }}`) is actually a column name in `df`, you may find `rlang::as_name()` and `rlang::ensym()` helpful for extracting the column name as a string.
:::

```{r}
# updated function with validation
```

---

## More complex functions

A big part of functions is that they actually use other functions to help out, this is often called a helper function. Since we already made a function to help us in filtering the table out we should make use of it as a helper function to make a more complex function.

While we are at it we should also add some default values to a function. Which just means that if a value isn't passed it will default to a preset value.

**4.Create a function that outputs yearly information related to a genre or actor**

The function we will now be making will collapse information that is filtered out from the previous function into rows based on year. For example if the previous function was used to filter horror genre movies, and it returned us 6 movies from 2011 we will want collapse the information into one row of 2011. To do this you would want to take the average rating and sum of votes.

While at it we should also introduce some default values to say if that year had a good, average, or bad quality of movies. So for this part if a year had an average rating above a 7 it should be considered good. Anything between a 4 and 7, including the 7 should be average, and anything below that should be bad. Although a user should be able to change those 7 and 4 values if they want

Here is an example of what the output should look like

| Genre | Year | avg_rating | total_votes| quality |
|:-----:|:----:|:----------:|:----------:|:-------:|
| Horror| 2011 |    7.1     | 104644     | good    |
| Horror| 2010 |    6.2     | 529576     | average |
| Horror| 2009 |    4       | 92455      | bad     |

: possible output

The input of this function should be straightforward, since your using a helper function you need its input to be included in this function as well. Then you are just adding the default variables that where mentioned before.

```{r}
# your function here
```

Test your function:

```{r}
yearly_info(movies_clean, Genre, "Horror") |> 
  head(5) |> 
  kable()
```

---

## The power of functions

As you might have noticed when writing your more complex function the true beauty of functions is how little you have to repeat code and how they just shorten everything. To get a more solid appreciation of them we will now use our more complex function in getting some fun data

**5. Make a table with yearly information on 3 genres**

Using the more complex function you made, save a dataframe that contains yearly information on 3 genres. It should also contain the title of the highest rated movie of that year for each genre, its duration, its rating and votes.

:::{.callout-tip}
## Expanding rows in a dataframe

It would be useful to look into functions that could help you **bind** rows together or something like `map_dfr`

:::

```{r}
# your code here
```

Take a moment to appreciate how little repeated code was needed here compared to doing this without functions.

---


## Presenting Data Nicely

**6. Make a polished summary table**

Using the dataframe you created in question 5, create a clean, readable table. To keep things manageable, only include rows from **2010 onward**, filter out any rows where `Year` is missing, and show only the top 10 rows by average rating.

Your table should:

- Show only the most relevant columns (genre, year, summary stats, best movie info)
- Have clean, human-readable column names — use `kable(col.names = ...)` to rename them rather than `rename()`
- Round any numeric values to two decimal places
- Include a descriptive caption

```{r}
# your code here
```

---

**7. Visualize average rating over time by genre**

Using the dataframe you created in question 5, create a line graph with `ggplot2` showing how the average yearly rating for each genre has changed over time.

Your plot should:

- Use `geom_line()` and `geom_point()` to show the trend
- Use `color` to distinguish genres
- Use `shape` on the points to reflect the `quality` label (`"good"`, `"average"`, `"bad"`)
- Have a clear title, axis labels, and legend
- Use a clean theme

```{r}
# your code here
```
