---
title: "Lab Solutions"
author: "Bernado Bernal, Tyler Chung, Marco DiPierro"
format:
  html: 
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---
## Functions and Dataframes

The goal of this lab is to get a solid handle on how to develop functions in hand with a dataframe. As doing so can greatly decrease the amount of repeated code needed to get similar output from looking through a dataframe multiple times at different points of time. Think about how you might have a dataframe of candy and you end with a table of chocolate candies with selected information, only to find out you have to make a similar table with hard candies instead. To make the same table twice but with one difference here would be a lot of repeated code. Luckily with a function that can be reduced to a single line in both cases along with the code to make the function.

The dataset we are using for this lab comes from [kaggle](https://www.kaggle.com/datasets/amanbarthwal/imdb-movies-data). The dataset has various information on 10,000 movies collected from *IMDB*. As such you can expect to find title, genre, and rating information from this dataset, along with some other information. Visit the link to get the whole data dictionary in order to find out what each column means, we will be using version two of the data. While the data is not overly complex it contains enough information to make a couple of functions that pull out or modify data in interesting ways.

## Setup
```{r}
#load in libraries
library(tidyverse)
library(knitr)
```

```{r}
#load in imdb-movies-dataset.csv
movies <- read_csv("imdb-movies-dataset.csv")

```

## Simple Functions

Lets start off by making a simple function.

##Data Cleaning
Before we can begin working with this dataset, we need to clean it up. Looking at the movies dataset you may notice some missing values, duplicate entries, and columns that won't be useful for our analysis.
Using one dplyr pipeline, clean the movies dataset by:

Removing any rows where Title, Genre, or Rating are missing
Removing any duplicate movies based on Title
Converting Year to an integer variable
Only keeping the columns we will use — Title, Year, Genre, Rating, Cast, Director, and Metascore

Assign your cleaned data to a new variable called movies_clean use this data going forward.
```{r}
movies_clean <- movies |>
  filter(
    !is.na(Title),
    !is.na(Genre),
    !is.na(Rating)
  ) |>
  distinct(Title, .keep_all = TRUE) |>
  rename(Duration = "Duration (min)") |>
  select(Title, Year, Certificate, Duration, Genre, Rating,
    Cast, Director, Metascore, Votes)
```

**Write a function to filter a table by a specific actor or genre**

Looking at the movies dataset it could be seen that both genre and cast contain a list of comma separated values, rather than only containing one value. For this function if we are given a specific actor or genre we should return the rows that contain that value, but with only the specified value given in the respective column. If given Horror as a genre then we should only return rows that contain Horror as a genre and mutate them to only contain Horror as a genre rather than something like "Horror, Comedy".

To help guide you in making this function you are taking 3 inputs.

1. A dataframe
2. A given column
3. A character value 

:::{.callout-tip}
## Working with dplyr in functions

Remember that when in a function it would be useful to use `rlang` in order to get tidy evaluation for a column passed to a function.

:::

```{r}
#code your function
single_actor_genre <- function(df, col, str) {
  df |> 
    filter(str_detect(string = {{col}},
                      pattern = str)) |> 
    mutate({{col}} := str_extract(string = {{col}},
                                 pattern = str))
}
```

Check you function
```{r}
#will likely have to change function name and input here to see if your code worked
single_actor_genre(movies, Genre, "Horror") |> 
  select(Title, Genre) |> 
  head(5) |> 
  kable()
```

## References

Markdown syntax guide used for links and list  
<https://www.markdownguide.org/basic-syntax/>

Quarto guide in including callout blocks  
<https://quarto.org/docs/authoring/callouts.html>
